Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-08 csf-2017/paper.tex
changed land and lor to text
Minor editing of the intro.

Compromising software security is often based on attacks that break programming
language properties relied on by software authors, such as control flow
integrity, local state encapsulation etc. Commodity processors offer little
support for defending against such attacks: they offer security primitives with
only coarse-grained memory protection and limited compartmentalization
scalability. As a result, defenses against attacks on control flow integrity and
local state encapsulation necessarily rely on techniques like machine code
rewriting (e.g.~\cite{wahbe_efficient_1993,abadi_control-flow_2005}) or
verification (e.g.~\cite{morrisett_system_1999}), virtual machines with a native
stack (e.g.~the JVM~\cite{lindholm_java_2014}) or
randomization~\cite{forrest_building_1997}. Such techniques for retrofitting
protection techniques to commodity hardware have disadvantages in terms of
performance, complexity and/or security.

\emph{Capability machines} are a type of processors, intended to remediate these
limitations by presenting a better security model at the hardware
level~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740,Carter:1994:HSF:195473.195579}.\dominique{todo:
  more?} \lars{can we add that there has been renewed interest and
  refer to crash-safe project ?}
 Such hardware tags every word (in the register file and in memory) to
enforce a strict separation between numbers and capabilities, a kind of pointers
carrying some kind of authority. Memory capabilities carry the authority to read
and/or write to a range of memory locations. Additionally, the details vary but
all capability machines offer some form of \emph{object capabilities}, which
represent the authority to invoke a piece of code without exposing the code's
encapsulated private state (e.g. the M-Machine's enter capabilities or CHERI's
sealed code/data pairs).

>>>>

Compromising software security is often based on attacks that break programming
language properties relied upon by software authors, such as control flow
integrity, local state encapsulation etc. Commodity processors offer little
support for defending against such attacks: they offer security primitives with
only coarse-grained memory protection and limited compartmentalization
scalability. As a result, defenses against attacks on control flow integrity and
local state encapsulation are either limited to only certain common forms of
attacks (leading to an attack-defense arms race) or rely on techniques like
machine code rewriting
(e.g.~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
verification (e.g.~\cite{morrisett_system_1999}), virtual machines with a native
stack (e.g.~the JVM~\cite{lindholm_java_2014}) or
randomization~\cite{forrest_building_1997}. The latter techniques essentially
emulate protection techniques on existing hardware, at the cost of performance,
system complexity and/or security.

\emph{Capability machines} are a type of processors, intended to remediate these
limitations by presenting a better security model at the hardware level. They
are based on old ideas
(e.g.~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
but have recently received renewd interest with particularly the CHERI project
proposing new ideas and tackling practical challenges like backwards
compatibility and realistic OS
support~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
machines tag every word (in the register file and in memory) to enforce a
strict separation between numbers and capabilities, a kind of pointers carrying
some kind of authority. Memory capabilities carry the authority to read and/or
write to a range of memory locations. Additionally, the details vary but all
capability machines offer some form of \emph{object capabilities}, which
represent the authority to invoke a piece of code without exposing the code's
encapsulated private state (e.g. the M-Machine's enter capabilities or CHERI's
sealed code/data pairs).
